Analytical choices for analyzing multidimensional behavior - Many analysts test hypotheses about human speech.

### 2021-03-10 RR_manuscript/Draft_RR.tex
:bug: - fix bugs and render pdf

In order to effectively accumulate knowledge, science needs (i) to produce data that can be replicated using the original methods and (ii) to arrive at robust conclusions substantiated by the data.
In recent coordinated efforts to replicate published findings, the scientific disciplines have uncovered surprisingly low success rates in replication attempts (e.g., Open Science Collaboration 2015; Camerer et al. 2018) leading to what is now referred to as the \emph{replication crisis}.
Beyond the difficulties of replicating scientific findings, a growing body of evidence suggests that the theoretical conclusions drawn from data are often variable even when researchers have access to reliable data (\highlight{REFS}).
The latter situation has been referred to as the \emph{inference crisis} (Rotello, Heit, and Dubé 2015; Jeffrey J. Starns et al. 2019b) and is, among other things, rooted in the inherent flexibility of data analysis (often referred to as researcher degrees of freedom: Simmons, Nelson, and Simonsohn 2011; Gelman and Loken 2014).
Data analysis involves many different steps, such as inspecting, organizing, transforming, and modeling the data, to name a few.
Along the way, different methodological and analytical choices need to be made, all of which may influence the final interpretation of the data.
These researcher degrees of freedom are both a blessing and a curse.

They are a blessing because they afford us the opportunity to look at nature from different angles, which, in turn, allows us to make important discoveries and generate new hypothesis (e.g., Box 1976; Tukey 1977; De Groot 2014).
They are a curse because idiosyncratic choices can lead to categorically different interpretations, which eventually find their way into the publication record where they are taken for granted (Simmons, Nelson, and Simonsohn 2011).
Recent projects have shown that the variability between different data analysts is vast.
This variability can lead independent researchers to draw different conclusions about the same data set as demonstrated by several projects crowd-sourcing analysis strategies (e.g., Silberzahn et al. 2018; Jeffrey J. Starns et al. 2019b; Botvinik-Nezer et al. 2020b).
These projects, however, might still underestimate the extent to which analysts vary because data analysis is not merely restricted to statistical inference.
Human behavior is complex and offers many ways to be translated into numbers.
This is particularly true for fields that draw conclusions about human behavior and cognition from multidimensional data like audio or video data.
In fields working on human speech production, for example, researchers need to make numerous decisions about what to measure and how to measure it.
This is not trivial, given the temporal extension of the acoustic signal and its complex structural composition.
Not only can decisions about measuring the signal influence downstream decisions about statistical modeling, but statistical results or modeling issues can also lead researchers to go back and revise earlier decisions about the measuring process itself.

In this article, we investigate the variability in analytic choices when many analyst teams analyze the same speech production data, a process that involves both decisions regarding the operationalization of a complex observed signal and decisions regarding the statistical modeling.
Specifically, we report the impact of the analytic pipeline on research results obtained by XX teams who gained access to the same set of acoustic recordings in order to answer the same research question.

@@

Data analysis comes with many decisions like how to measure a given phenomenon or behavior, what data to submit to statistical modeling and which to exclude in the final analysis, or what inferential decision procedure to apply.
\highlight[id=JC, comment = {I'd probably drop this. When we make unspecified decisions during data analysis we can stumble upon spurious results (most likely) as well as real results (less likely, still possible)... in other words it might not all be error... the issue is that we won't know the difference}]{However, if these decisions} during data analysis are not specified in advance, we might stumble upon seemingly meaningful patterns in the data that are merely statistical flukes.\comment[id=TR]{But isn't the human bias the major issue in statistical misinterpretation? (It also leads naturally to the next paragraph.)}
This can be problematic because humans show cognitive biases that can lead to erroneous inferences.
Humans filter the world in irrational ways (e.g., Tversky and Kahneman 1974), seeing coherent patterns in randomness (Brugger 2001), convincing themselves of the validity of prior expectations ({``I knew it,''} Nickerson 1998), and perceiving events as being plausible in hindsight ({``I knew it all along,''} Fischhoff 1975).
\comment[id=JC]{If helpful we could also give a concrete common example here, like regarding forecasting the weather or something}
\comment[id=TR]{Generally a good idea, but I'd say we keep it crisp at that point. The take-home here is: Humans are biased.}
In connection with an academic incentive system that rewards certain discovery processes more than others (Sterling 1959; Koole and Lakens 2012), we often find ourselves exploring many possible analytical pipelines, but only reporting a select few.
This issue is particularly amplified in fields in which the raw data lend themselves to many possible ways to measure (Roettger 2019).
Combined with a wide variety of methodological and theoretical traditions as well as varying levels of statistical training across subfields, the inherent flexibility of data analysis might lead to a vast plurality of analytic approaches that can lead to different scientific conclusions.
Consequently, there might be many published papers that present overconfident interpretations of their data based on idiosyncratic analytic strategies (e.g., Simmons, Nelson, and Simonsohn 2011; Gelman and Loken 2014).
These interpretations are either associated with an unknown amount of uncertainty or lend themselves to alternative interpretation if analyzed differently.
However, instead of being critically evaluated, scientific results often remain unchallenged in the publication record.
Despite recent efforts to improve transparency and reproducibility (REFS) and freely available and accessible infrastructures such as provided by the Open Science Framework (osf.io, ADD), critical re-analyses of published analytic strategies are still not very common because data sharing remains rare (Jelte M. Wicherts et al. 2006).

While this issue has been widely discussed both from a conceptual point of view (Simmons, Nelson, and Simonsohn 2011; Wagenmakers et al. 2012; Nosek and Lakens 2014) and its application in individual scientific fields (e.g.~Wichert et al.~2015, Charles et al.~2019, Roettger 2019), there are still many unknowns regarding the extent of analytical plurality in practice.
Recent collaborative attempts have started to shed light on how different analysts tackle the same data set and have revealed a large amount of variability.

@@

In a collaborative effort, Silberzahn et al. (2018) let twenty-nine independent analysis teams address the same research hypothesis.
Analytical approaches and consequently the results varied widely between teams.
Sixty-nine percent of the teams found support for the hypothesis, and 31\% did not.
Out of the 29 analytical strategies, there were 21 unique combinations of covariates.
Importantly, the observed variability was neither predicted by the team's preconceptions about the phenomenon under investigation nor by peer ratings of the quality of their analyses.
The authors results suggest that analytic plurality is a fact of life and not driven by different levels of expertise or bias.

Similar crowd-sourced studies recruiting independent analyst teams showed similar results.
Both Dutilh et al. (2019) ad Jeffrey J. Starns et al. (2019a) investigated analysts choices when inferring theoretical constructs based on the same data set using computational models.
Both studies reveal vastly different modeling strategies, even though scientific conclusions were similar across analysis teams (see also Parker et al.~(accepted) on analytical flexibility in ecology and Botvinik-Nezer et al. 2020a on neuroimaging data).
Bastiaansen et al. (2020) crowdsourced clinical recommendations based on data analysis of an individual patient.
Results suggests that analysis team differed substantially regarding decisions related to both the statistical analysis of the data set and the theoretical rationale behind interpreting the statistical results, indicating that there are multiple layers for degrees of freedom that interact with each other.

While these projects show a large degree of analytical flexibility with impactful consequences, they dealt with flexibility in inferential or computational modeling.
In these studies the data sets were fixed and data collection or measurement could not be changed.\comment[id=JC]{any particular reason to favor data table over data set?}\comment[id=TR]{I think my reasoning was data set can be data table or raw data. Data table is the numerical derivation of the raw data. Maybe not a useful distinction?}

However, in many fields the primary raw data are complex signals that need to be operationalized in light of the research question.
This is especially true in the social sciences, where the subject of investigation corresponds to observable and unobservable human behavior.
Decisions about how to measure a theoretical construct related to such behavior or its underlying cognitive processes might interact with downstream decisions about statistical modeling and vice versa.
For example, Flake and Fried (2020) discuss the cascading impact different practices can have on several components of psychometric research.
The authors highlight, among others, the following degrees of freedom in the choice and development of measures: definition of the theoretical construct, justification of the selected measure, description of the measure and of how it maps onto the construct, response coding and related transformations, and post-hoc modifications to the chosen measure.
Taken together, these aspects alone exponentially increase the combinations of possible analytical choices, and hence flexibility in research outcomes.

In those social sciences concerned with communication, human behaviour is sometimes measured as a complex visual or acoustic signal.
This added complexity introduces further fork junctions in the research pipe-line, thus increasing yet again analytical flexibility.
The present study looks at experimentally elicited speech data to understand how this flexibility manifests itself in a scenario where complex decision procedures are involved in the operationalization and measuring of complex signals.

@@

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\tightlist
\item
  ``I can't bear another meeting on Zoom.''
\end{enumerate}

Depending on the speaker's intention, this sentence can be said in different ways.
If, for instance, the speaker is exhausted by all their meetings, the speaker might acoustically highlight the word \emph{another} or \emph{meeting}.
If, on the other hand, the speaker is just tired of video conferences, they might acoustically highlight the word \emph{Zoom}.

If we decide to compare the speech signal associated with these two intentions, how can we quantify the difference between them?
Given their physical manifestation (speech), what do we measure and how do we measure it?
Because of the continuous and transient nature of speech, identifying speech parameters and temporal domains within which to measure those parameters becomes a non-trivial task.
Utterances stretch over hundreds of milliseconds and contain several levels of linguistically relevant units such as phrases, words, syllables, and individual sounds.
The researcher is thus confronted with a considerable number of parameters and combinations thereof to choose from.

Speech categories are inherently multidimensional and dynamic: they consist of a cluster of parameters that are modulated over time.
The acoustic parameters of one category are usually asynchronous, i.e.~they appear at different time points in the unfolding signal, and overlap with parameters of other categories {[}e.g., Jongman, Wayland, and Wong (2000); Lisker (1986); Summerfield, 1984; Winter (2014){]}.
A classical example is the distinction between voiced and voiceless stops in English (i.e.~/b/ and /p/ in \emph{bear} vs \emph{pear}).
This voiced/voiceless contrast is manifested by many acoustic features which can differ depending on several factors, such as position of the consonant in the word and surrounding sounds (Lisker 1977).
Furthermore, correlates of the contrast can even be found away from the consonant, in temporally distant speech units.
For example, the initial /l/ of the English words \emph{led} and \emph{let} is affected by the voicing of the final consonant (/t, d/) (Hawkins and Nguyen 2004).

The multiplicity of phonetic cues grows exponentially if we look at larger temporal domains as is the case for suprasegmental aspects of speech.
For example, studies investigating acoustic correlates of word stress (e.g.~the difference between \emph{insíght} and \emph{íncite}) have been using a wide variety of measurements, including temporal characteristics (duration of certain segments or sub-segmental intervals), spectral characteristics (intensity, formants, and spectral tilt), and measurements related to fundamental frequency (f0) (e.g., Gordon and Roettger 2017).
Moving onto the expression of higher-level functions like information structure and discourse pragmatics, relevant acoustic cues can be distributed throughout even larger domains, such as phrases and whole utterances (e.g., Ladd 2008).
Differences in position, shape, and alignment of pitch modulations over multiple locations within a sentence are correlated with differences in discourse functions (e.g.~Niebuhr et al., 2011).
The latter can also be expressed by global vs local pitch modulations (Van Heuven et al. 2002), as well as acoustic information within the temporal or spectral domain (e.g., Van Heuven and Van Zanten 2005).
Extra-linguistic information, like speaker's intentions, levels of arousal or social identity, are also conveyed by broad-domain parameters, such as voice quality, rhythm, and pitch (Foulkes and Docherty 2006; Ogden 2004; White, Payne, and Mattys 2009).

When testing hypotheses on speech production data, researchers are faced with many choices and possibilities.
The larger the functional domain (e.g.~segments vs words vs utterances), the higher the number of conceivable operationalizations.
For example, several decisions have to be made when comparing the two realization of the sentence in (1), one of which is intended to signal emphasis on \emph{another} and one of which emphasizes \emph{zoom}.

@@

Do we compare only the word \emph{another} in (2a) and (2b), or also the word \emph{zoom}?
Do we measure utterance-wide acoustic profiles, whole words, or just the stressed syllables?
Do we average across the chosen time domain or do we measure a specific point in time?
Do we measure fundamental frequency, intensity, something else?

When looking at phrase-level temporal domains, the number of possible analytical pipelines quickly explodes.
This plurality of analytical paths is illustrated in Figure \ref{fig:forkingPaths}.
When comparing two utterance such as (2a) and (2b), there are many things to consider.
Even if we have decided to compare fundamental frequency (f0) of only the word \emph{another} across the two utterances, there are still many choices to be made, all of which need to be justified.
For example, we could measure f0 at specific points in time like the onset of the window, the offset, or the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarise f0 across the entire window and extract the mean, median or standard deviation of f0.
And the garden of forking paths does not stop here.
In \highlight[id=SC, comment = {The figure text is quite small. Probably better to change to a longer rather then wider layout.}]{Figure} \ref{fig:forkingPaths}, we went with a specific option to automatically calculate f0, INSERT SOME EXAMPLES OF PITCH TRACKING OPTIONS.\comment{add examples of pitch tracking options}
Moreover, knowing that the signals and the measures thereof can be noisy, we could smooth these using one algorithm and degree over others, or remove values that substantially deviate from surrounding values or expectations, either manually or automatically, and so on.

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/ste/repos/many_analyses/figs/ForkingPaths} \caption{Illustrations of the analytical flexibility associated with acoustic analyses. (A) an example of multiple possible and justifiable decisions when comparing to utterances; (B) waveform and fundamental frequency (f0) track of two instances of utterances 2a and 2b. The words ``another is highlighted by the green box''; (C) spectrogram and f0 track of the word another, exemplifying different operationalizations of differences in pitch.}\label{fig:forkingPaths}
\end{figure}

These decisions are usually made prior to any statistical analysis, but are at times revised a posteriori (i.e.~after data collection and/or preliminary analyses) in light of unforeseen or surprising outcomes.
These myriads of possible decisions are exponentiated by researcher degrees of freedom related to statistical analysis (e.g. Jelte M. Wicherts et al. 2016).
Even the analysis of a single measure can be approached via an ever-increasing range of different statistical models (\highlight{REFs}).
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines across independent researchers, we provided XX analytical teams with an experimentally elicited speech corpus and asked them to investigate acoustic differences related to a functional contrast that might be manifested across the whole utterance.\comment[id=SC]{This needs a smother transition.}

\hypertarget{the-data-set-the-acoustic-properties-of-redundant-modifiers}{%
\subsection{The data set: The acoustic properties of redundant modifiers}\label{the-data-set-the-acoustic-properties-of-redundant-modifiers}}

The data set given to the analysis teams comes from a study that investigated whether speakers acoustically modify utterances to signal typicality of noun/adjective combinations in referring expressions such as \emph{a blue banana} (atypical) vs \emph{a yellow banana} (typical).
Referring is one of the most basic and prevalent uses of language and one of the most widely researched areas in language science.
It is an open question how speakers choose a referring expression when they want to refer to a specific entity like a banana.
The context within which an entity occurs (i.e., with other non-fruits, other fruits, or other bananas) plays a large part in determining the choice of referring expressions.
Generally, speakers aim to be as informative as possible to uniquely establish reference to the intended object, but they are also resource-efficient in that they avoid redundancy (Grice 1975).
Thus one would expect the use of a modifier, for example, only if it is necessary for disambiguation.
For instance, one might use the adjective \emph{yellow} to describe a banana in a situation in which there is a yellow and a less ripe green banana available, but not when there is only one banana to begin with.

Despite this coherent idea that speakers are both rational and efficient, there is much evidence that speakers are often over-informative: Speakers use referring expressions that are more specific than strictly necessary for the unambiguous identification of the intended referent (Sedivy 2003; Westerbeek, Koolen, and Maes 2015; Rubio-Fernández 2016), which has been argued to facilitate object identification and making communication between speakers and listeners more efficient (Arts et al. 2011; Paraboni, Van Deemter, and Masthoff 2007; Rubio-Fernández 2016).
Recent findings suggest that the utility of a referring expression depends on how good it is for a listener (compared to other referring expressions) to identify a target object.
For example, Degen et al. (2020) showed that modifiers that are less typical for a given referent (e.g.~a blue banana) are more likely to be used in an over-informative scenario (e.g.~when there is just one banana).
This account, however, has mainly focused on content selection (Gatt et al. 2013), i.e.~whether a certain referential expression is chosen or not, ignoring the fact that speech communication is much richer.

Even looking at morphosyntactically identical expressions, speakers can modulate these via suprasegmental acoustic properties like temporal and spectral modifications of the segments involved (e.g., Ladd 2008).
Most prominently, languages use intonation to signal discourse relationships between referents (among other functions).
Intonation marks discourse-relevant referents for being new or given information to guide the listeners' interpretation of incoming messages.
In many languages, speakers can use particular pitch movements to signal whether a referent has already been mentioned and is therefore referred back to, or a referent is newly introduced into the discourse.
Many languages use intonation in order to signal if a referent is contrasting with one or more alternatives that are relevant to the current discourse.
Content selection aside, in a scenario in which a speaker wants to refer to a banana when there is also a pear on the table, the speaker would most likely produce a high rising pitch accent on \emph{banana} to indicate the contrastive nature of the noun.
In a scenario in which the speaker wants to refer to a yellow banana when there is also a less ripe green banana on the table, the speaker would most likely produce a high rising pitch accent on \emph{yellow} to indicate the contrastive nature of the modifier.
In addition to a pitch accent, elements that are new and/or contrastive are often produced with additional suprasegmental prominence, i.e.~segments are hyperarticulated, resulting in longer, louder and more clearly articulated acoustic targets.

To answer the question of whether speakers modify speech to signal atypical referents, thirty native German speakers were recorded in a speech production study while interacting with a confederate (one of the experimenters).
The participants had to verbally instruct the confederate to select a specified target object out of four object presented on a screen.
The subject and confederate were seated at the opposite sides of a table, each facing one of two mirrored computer screens.
The participant and the experimenter could not see each other nor each others' screens.
After a familiarisation phase (see XXX), the subject first saw four colored objects in the top left, top right, bottom left, and bottom right corners of the screen.
One of the objects served as the target, another as the competitor, and the remaining two objects served as distractors.
Objects were referred to using noun phrases consisting of an adjective modifier denoting color and a modified object (e.g.~\emph{Gelbe Zitrone} `yellow lemon,' \emph{Rote Gurke} `red cucumber,' \emph{Rote Socken} `red socks').

>>>>

In order to effectively accumulate knowledge, science needs to (i) produce data that can be replicated using the original methods and (ii) arrive at robust conclusions substantiated by the data.
In recent coordinated efforts to replicate published findings, scientific disciplines have uncovered surprisingly low success rates (e.g., Collaboration 2015; Camerer et al. 2018) leading to what is now referred to as the \emph{replication crisis}.
Beyond the difficulties of replicating scientific findings, a growing body of evidence suggests that the theoretical conclusions drawn from data are often variable even when researchers have access to reliable data.
The latter situation has been referred to as the \emph{inference crisis} (Rotello, Heit, and Dubé 2015; Starns et al. 2019) and is, among other things, rooted in the inherent flexibility of data analysis (often referred to as researcher degrees of freedom: Simmons, Nelson, and Simonsohn 2011; Gelman and Loken 2014).
Data analysis involves many different steps, such as inspecting, organizing, transforming, and modeling the data, to name a few.
Along the way, different methodological and analytical choices need to be made, all of which may influence the final interpretation of the data.
These researcher degrees of freedom are both a blessing and a curse.

They are a blessing because they afford us the opportunity to look at nature from different angles, which, in turn, allows us to make important discoveries and generate new hypothesis (e.g., Box 1976; Tukey 1977; De Groot 2014).
They are a curse because idiosyncratic choices can lead to categorically different interpretations, which eventually find their way into the publication record where they are taken for granted (Simmons, Nelson, and Simonsohn 2011).
Recent projects have shown that the variability between different data analysts is vast and can lead independent researchers to draw different conclusions from the same data set (e.g., Silberzahn et al. 2018; Starns et al. 2019; Botvinik-Nezer et al. 2020b).
These studies, however, might still underestimate the extent to which analysts vary because data analysis is not merely restricted to statistical inference.
Human behavior is complex and offers many ways to be translated into numbers.
This is particularly true for fields that draw conclusions about human behavior and cognition from multidimensional data like audio or video data.
In fields working on human speech production, for example, researchers need to make numerous decisions about what to measure and how to measure it.
This is not trivial, given the temporal extension of the acoustic signal and its complex structural composition.
Not only can decisions about measuring the signal influence downstream decisions about statistical modeling, but statistical results or modeling issues can also lead researchers to go back and revise earlier decisions about the measuring process itself.

In this article, we will investigate the variability in analytic choices when many analyst teams analyze the same speech production data set, a process that involves both decisions regarding the operationalization of a complex signal and decisions regarding the statistical analysis.
Specifically, we will report the impact of analytic decision making on research results obtained by a minimum of 12 teams who will gain access to the same set of acoustic recordings in order to answer the same research question.

@@

Data analysis comes with many decisions, like how to measure a given phenomenon or behavior, which data to submit to statistical modeling and which to exclude in the final analysis, or what inferential decision making procedure to apply.
This can be problematic because humans show cognitive biases that can lead to erroneous inferences.
Humans filter the world in irrational ways (e.g., Tversky and Kahneman 1974), seeing coherent patterns in randomness (Brugger 2001), convincing themselves of the validity of prior expectations ({``I knew it,''} Nickerson 1998), and perceiving events as being plausible in hindsight ({``I knew it all along,''} Fischhoff 1975).
In conjunction with an academic incentive system that rewards certain discovery processes more than others (Sterling 1959; Koole and Lakens 2012), we often find ourselves exploring many possible analytical pipelines, but only reporting a select few.
This issue is particularly amplified in fields in which the raw data lend themselves to many possible ways of being measured (Roettger 2019).
Combined with a wide variety of methodological and theoretical traditions as well as varying levels of statistical training across subfields, the inherent flexibility of data analysis might lead to a vast plurality of analytic approaches that can lead to different scientific conclusions.
Consequently, there might be many published papers that present overconfident interpretations of their data based on idiosyncratic analytic strategies (e.g., Simmons, Nelson, and Simonsohn 2011; Gelman and Loken 2014).
These interpretations are either associated with an unknown amount of uncertainty or lend themselves to alternative interpretation if analyzed differently.
However, instead of being critically evaluated, scientific results often remain unchallenged in the publication record.
Despite recent efforts to improve transparency and reproducibility (e.g. Miguel et al. 2014; Klein et al. 2018) and the advent of freely available and accessible infrastructures, such as those provided by the Open Science Framework (osf.io), critical re-analyses of published analytic strategies are still uncommon because data sharing remains rare (Wicherts et al. 2006).

Analytical flexibility has been widely discussed from a conceptual point of view (Simmons, Nelson, and Simonsohn 2011; Wagenmakers et al. 2012; Nosek and Lakens 2014) and in regard to its application in individual scientific fields (e.g.~Wichert et al.~2015, Charles et al.~2019, Roettger 2019), notwithstanding, there are still many unknowns regarding the extent of analytical plurality in practice.
Recent collaborative attempts have started to shed light on how different analysts tackle the same data set and have revealed a large amount of variability.

@@

In a collaborative effort, Silberzahn et al. (2018) let twenty-nine independent analysis teams address the same research hypothesis.
Analytical approaches and consequently the results varied widely between teams.
Sixty-nine percent of the teams found support for the hypothesis, and 31\% did not.
Out of the 29 analytical strategies, there were 21 unique combinations of covariates.
Importantly, the observed variability was neither predicted by the team's preconceptions about the phenomenon under investigation nor by peer ratings of the quality of their analyses.
The authors results suggest that analytic plurality may be an inevitable byproduct of the scientific process and not necessarily driven by different levels of expertise nor bias.

Several other recent studies corroborated this analytical flexibility across different disciplines.
Both Dutilh et al. (2019) ad Starns et al. (2019) investigated analysts' choices when inferring theoretical constructs based on the same data set using computational models.
Both studies revealed vastly different modeling strategies, even though scientific conclusions were similar across analysis teams (see also Parker et al.~(accepted) on analytical flexibility in ecology and Botvinik-Nezer et al. 2020a on neuroimaging data).
Bastiaansen et al. (2020) crowdsourced clinical recommendations based on analyses of an individual patient.
Results suggests that analysts differed substantially regarding decisions related to both the statistical analysis of the data and the theoretical rationale behind interpreting the statistical results.

While these studies attested a large degree of analytical flexibility with possibly impactful consequences, they only investigated analytical decisions related to inferential statistics or the architecture of computational models.
In these studies the data sets were fixed and neither data collection nor measurement could be changed.
However, in many fields the primary raw data are complex signals that need to be operationalized relative to a theoretically motivated research question.
This is especially true in the social sciences, where the phenomenon under investigation corresponds to both observable and unobservable human behavior.
Decisions about how to measure a theoretical construct related to such behavior or its underlying cognitive processes might interact with downstream decisions about statistical modeling and vice versa.

For instance, Flake and Fried (2020) discuss the cascading impact that different practices can have on psychometric research.
The authors highlight, among others, the following degrees of freedom in the choice and development of measures: definition of the theoretical construct, justification of the selected measure, description of the measure and of how it maps onto the construct, response coding and related transformations, and post-hoc modifications to the chosen measure.
Taken together, these aspects alone exponentially increase the combinations of possible analytical choices, and hence flexibility in research outcomes.

In social sciences concerned with communication, human behavior is sometimes measured as a complex visual or acoustic signal.
This added complexity introduces further fork junctions in the research pipe-line, and in turn further increases analytical flexibility.
The present study looks at experimentally elicited speech production data in order to understand how this flexibility manifests itself in a scenario where complex decision procedures are involved in the operationalization and measuring of complex signals.

@@

\vspace{1em}

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\tightlist
\item
  ``I can't bear another meeting on Zoom.''
  \vspace{1em}
\end{enumerate}

Depending on the speaker's intention, this sentence can be said in different ways.
If, for instance, the speaker is exhausted by all their meetings, the speaker might acoustically highlight the word \emph{another} or \emph{meeting} to contrast it with other, more pleasant activities.
If, on the other hand, the speaker is just tired of video conferences, as opposed to say face-to-face meetings, they might acoustically highlight the word \emph{Zoom}.

If we decide to compare the speech signal associated with these two intentions, how can we quantify the difference between them?
Given their physical manifestation (speech), what do we measure and how do we measure it?
Because of the continuous and transient nature of speech, identifying speech parameters and temporal domains within which to measure those parameters becomes a non-trivial task.
Utterances stretch over several thousand milliseconds and contain different levels of linguistically relevant units such as phrases, words, syllables, and individual sounds.
The researcher is thus confronted with a considerable number of parameters and combinations thereof to choose from.

Speech categories are inherently multidimensional and dynamic: they consist of clusters of parameters that are modulated over time.
The acoustic parameters of one category are usually asynchronous, i.e.~they appear at different time points in the unfolding signal, and overlap with parameters of other categories (e.g. Jongman, Wayland, and Wong 2000; Lisker 1986; \textbf{summerfield1984?}; Winter 2014).
A classical example is the distinction between voiced and voiceless stops in English (i.e.~/b/ and /p/ in \emph{bear} vs \emph{pear}).
This lexical contrast is manifested by many acoustic features which can differ depending on several factors, such as position of the consonant in the word and surrounding sounds (Lisker 1977).
Furthermore, correlates of the contrast can even be found away from the consonant, in temporally distant speech units.
For example, the initial /l/ of the English words \emph{led} and \emph{let} is affected by the voicing of the final consonant (/t, d/) (Hawkins and Nguyen 2004).

The multiplicity of phonetic cues grows exponentially if we look at larger temporal domains as is the case for suprasegmental aspects of speech.
For example, studies investigating acoustic correlates of word stress (e.g.~the difference between \emph{ínsight} and \emph{incíte}) have been using a wide variety of measurements, including temporal characteristics (duration of certain segments or sub-segmental intervals), spectral characteristics (intensity, formants, and spectral tilt), and measurements related to fundamental frequency (f0) (e.g., Gordon and Roettger 2017).
Moving on to the expression of higher-level functions, like information structure and discourse pragmatics, relevant acoustic cues can be distributed throughout even larger domains, such as phrases and whole utterances (e.g., Ladd 2008).
Differences in position, shape, and alignment of pitch modulations over multiple locations within a sentence are correlated with differences in discourse functions (e.g.~Niebuhr et al., 2011).
The latter can also be expressed by global vs local pitch modulations (Van Heuven et al. 2002), as well as acoustic information within the temporal or spectral domain (e.g., Van Heuven and Van Zanten 2005).
Extra-linguistic information, like the speaker's intentions, levels of arousal or social identity, are also conveyed by broad-domain parameters, such as voice quality, rhythm, and pitch (Foulkes and Docherty 2006; Ogden 2004; White, Payne, and Mattys 2009).

In short, when testing hypotheses on speech production data, researchers are faced with many choices and possibilities.
The larger the functional domain (e.g.~segments vs words vs utterances), the higher the number of conceivable operationalizations.
For example, several decisions have to be made when comparing the two realization of the sentence in (1), one of which is intended to signal emphasis on \emph{another} and one of which emphasizes \emph{zoom}.

\vspace{1em}

@@

\vspace{1em}

Do we compare only the word \emph{another} in (2a) and (2b), or also the word \emph{zoom}?
Do we measure utterance-wide acoustic profiles, whole words, or just the stressed syllables?
Do we average across the chosen time domain or do we measure a specific point in time?
Do we measure fundamental frequency, intensity, or something else?

When looking at phrase-level temporal domains, the number of possible analytical pipelines increases exponentially.
This plurality of analytical paths is illustrated in Figure \ref{fig:forkingPaths}.
Even if we have decided to compare fundamental frequency (f0) of only the word \emph{another} across the two utterances, there are still many choices to be made, all of which need to be justified.
For example, we could measure f0 at specific points in time like the onset of the window, the offset, or the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarize f0 across the entire window and extract the mean, median or standard deviation of f0, all of which have been used to analyze speech data (see Gordon and Roettger 2017).
And the garden of forking paths does not stop here.
Figure \ref{fig:forkingPaths}, illustrates a specific option to automatically calculate f0.
Knowing that the signals and the measures thereof can be noisy, another option would be to smooth them using one algorithm and degree over others, or remove values that substantially deviate from surrounding values or expectations, either manually or automatically, and so on.

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/ste/repos/many_analyses/figs/ForkingPaths} \caption{Illustrations of the analytical flexibility associated with acoustic analyses. (A) an example of multiple possible and justifiable decisions when comparing two utterances; (B) waveform and fundamental frequency (f0) track of two instances of utterances 2a and 2b. The words ``another is highlighted by the green box''; (C) spectrogram and f0 track of the word another, exemplifying different operationalizations of differences in pitch.}\label{fig:forkingPaths}
\end{figure}

These decisions are usually made prior to any statistical analysis, but are at times revised a posteriori (i.e.~after data collection and/or preliminary analyses) in light of unforeseen or surprising outcomes.
These myriads of possible decisions are exponentiated by those researcher degrees of freedom related to statistical analysis (e.g. Wicherts et al. 2016).
In sum, speech data are complex physical signals that constitutes an as of yet unappreciated amount of analytical flexibility when choosing measures and operationalizing underlying theoretical constructs.
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines across independent researchers, we will provide analytical teams with an experimentally elicited speech production data set.
The study aimed at investigating whether speakers acoustically modify utterances to signal unexpected referring expressions.
In the following, we will shortly introduce the research question of the original study.

\hypertarget{the-data-set-the-acoustic-properties-of-atypical-modifiers}{%
\subsection{The data set: The acoustic properties of atypical modifiers}\label{the-data-set-the-acoustic-properties-of-atypical-modifiers}}

Referring is one of the most basic and prevalent uses of language and one of the most widely researched areas in language science.
It is an open question how speakers choose a referring expression when they want to refer to a specific entity like a banana.
The context within which an entity occurs (i.e., with other non-fruits, other fruits, or other bananas) plays a large part in determining the choice of referring expressions.
Generally, speakers aim to be as informative as possible to uniquely establish reference to the intended object, but they are also resource-efficient in that they avoid redundancy (Grice 1975).
Thus one would expect the use of a modifier, for example, only if it is necessary for disambiguation.
For instance, one might use the adjective \emph{yellow} to describe a banana in a situation in which there is a yellow and a less ripe green banana available, but not when there is only one banana to begin with.

Despite the coherent idea that speakers are both rational and efficient, there is much evidence that speakers are often over-informative: Speakers use referring expressions that are more specific than strictly necessary for the unambiguous identification of the intended referent (Sedivy 2003; Westerbeek, Koolen, and Maes 2015; Rubio-Fernández 2016), which has been argued to facilitate object identification and making communication between speakers and listeners more efficient (Arts et al. 2011; Paraboni, Van Deemter, and Masthoff 2007; Rubio-Fernández 2016).
Recent findings suggest that the utility of a referring expression depends on how good it is for a listener (compared to other referring expressions) to identify a target object.
For example, Degen et al. (2020) showed that modifiers that are less typical for a given referent (e.g.~a blue banana) are more likely to be used in an over-informative scenario (e.g.~when there is just one banana).
This account, however, has mainly focused on content selection (Gatt et al. 2013), i.e.~whether a certain referential expression is chosen or not, ignoring the fact that speech communication is much richer.

Even looking at morphosyntactically identical expressions, speakers can modulate utterances via suprasegmental acoustic properties like temporal and spectral modifications of the segments involved (e.g., Ladd 2008).
Most prominently, languages use intonation to signal discourse relationships between referents (among other functions).
Intonation marks discourse-relevant referents for being new or given information to guide the listeners' interpretation of incoming messages.
Beyond structuring information relative to the discourse, a few studies suggested that speakers might use intonation to signal atypical lexical combinations {[}e.g.~dimitrova2008prosodic; dimitrova2009did{]}.
Referential expressions such as \emph{blue banana} were produced with greater prosodic prominence than more typical referent such as \emph{yellow banana}.
If true, these findings are in line with the idea of resource-efficient, rational language users who modulate their speech in order to facilitate listeners' comprehension process.
However, the original conclusions of above studies were based on small (n=10), unjustified sample sizes.
Moreover, the data is not available, and the statistical analysis used anti-conservative co-variance structures, leaving reason to doubt the original conclusions.

To further illuminate the question of whether speakers modify speech to signal atypical referents, we recorded thirty native German speakers interacting with a confederate (one of the experimenters) in a referential game.
The participants had to verbally instruct the confederate to select a specified target object out of four object presented on a screen.
The subject and confederate were seated at the opposite sides of a table, each facing one of two mirrored computer screens.
The participant and the experimenter could not see each other nor each others' screens.
After a familiarization phase (see XXX), the subject first saw four colored objects in the top left, top right, bottom left, and bottom right corners of the screen.

One of the objects served as the target, another as the competitor, and the remaining two objects served as distractors.
Objects were referred to using noun phrases consisting of an adjective modifier denoting color and a modified object (e.g.~\emph{Gelbe Zitrone} `yellow lemon,' \emph{Rote Gurke} `red cucumber,' \emph{Rote Socken} `red socks').

In the center of the screen, a black cube was displayed, which could be moved by the experimenter.
The participant would read a sentence prompt out loud (\emph{Du sollst den Würfel auf der ablegen} `You have to put the cube on top of the ') to instruct the experimenter to drag the cube on top of one of the four depicted objects (the \emph{competitor}) using the mouse.
After the experimenter had moved the cube as instructed, the subject would read another sentence prompt (\emph{Und jetzt sollst du den Würfel auf der ablegen} `And now, you have to put the cube on top of the ') instructing the experimenter to move the cube on top of a different object (the \emph{target}).
The second utterance in the trial was the critical trial for analysis.
